Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-07 how.tex
Updates from ShareLaTeX

 \documentclass[conference]{IEEEtran}
\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 %\usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}

\usepackage[running]{lineno}
\usepackage{program}
\usepackage{cite}
\usepackage{alltt}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\definecolor{lightgray}{gray}{0.975}
\usepackage{fancyvrb}
\usepackage{stfloats}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amsmath} 
\usepackage[caption=false,font=footnotesize]{subfig}
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=none,
numberstyle=\scriptsize,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\scriptsize\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
%
}}{}

@@

Data can be rapidly and recursively divided   as follows.
Find   two   distance instances,  $X,Y$
by picking any instance $W$ at random, then setting $X$ to its most
distant instance, then setting $Y$ to the instance most distant from
$X$~\cite{fastmap}
(this requires only $O(2N)$ comparisons
of $N$ instances).
Next, project each instance $Z$
onto a line that  runs between $X,Y$ using the cosine
rule of \fig{where}b. After that,  split the data at the median $x$ value of all instances and
recurses on each half  (stopping when
one half has less  than $\sqrt{N}$ of the original population). See \fig{code1d} for further details.

@@

Let   $X,Y$ be two points joined by  a line  length $c$.
A third point, $Z$, has distances  $a=dist(Z,X)$ and
$b=dist(Z,Y)$ to $X,Y$, respectively.
According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
Further, according to Pythagoras' theorem, $Z$ stands at a distance
$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$. See \fig{code1c} for further details.

@@

HERE's feature weighting algorithm
comes from the CART regression tree learner~\cite{Breiman1984}.
It sorts independent variables
 according to how well they can reduce the variance
of a  numeric objective.
If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
then the expected
value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
where $v$ is the standard deviation of the objective.
A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
then split at $\argmin_i w_i$, then recursing into each half.  See \fig{code1d} for further details

@@

Those same regions can be used for instance-based 
{\em planning}-- but instead of leaving test instances in the regions
found during training, we {\em displace} them towards regions
with better objective scores.
To achieve this, 
during training, HOW clusters data using the procedure of \fig{where}a,
then draws lines between the centroids of  clusters $X,Y$ that are nearest
neighbors.
By convention, we say that the cluster containing $X$ has the  better objective scores. This means we can say that the line is a useful direction for displacing
data; i.e.  away from $Y$ and towards $X$.
Hence,  during testing,  HOW use Pythagoras' theorem
and \fig{where}c to find the  nearest line to every test instance. That test instance is   then displaced towards the better end
of that line (towards  $X$).

An important design decision with HOW 
is how to determine the exact {\em direction} and {\em magnitude} of the
displacement.  Following on from the discussion around   ISSUE~\#1,
we note that a design goal of HOW is
{\em minimal displacements} of project data.
To implement minimal displacement:
\bi
\item
HOW sorts  the independent variables
using the procedure of  \fig{where}d. It then only
displaces the    best $B$\%  of that sort (while leaving the rest untouched).
In  the experiments shown below,  we explore   $B \in \{25,50,100\}\%$.
\item
HOW also  limits displacement
to very small regions in the space of the training data (never more that the separation
of the two closest clusters).  HOW's clusterer divides  data into many   small clusters containing
$\sqrt{N}$ of the population.
HOW   limits displacement to  \mbox{$0 \le F \le c$}, where
$c$ is the distance between the two closest clusters. 
In  the experiments shown below,  we explore
\mbox{$F\in \{0.25, 0.5, 0.75\}$}.
\ei
Turning now to the displacement {\em direction},
our experiments show that it useful to  carefully reflect on exactly where HOW should  push each test instance.
HOW replaces each cluster   with one {\em exemplar} that best
characterizes the data in that cluster. It then displaces the
test data towards that exemplar, as follows.
Let $Z_j$ be an independent variable within the
test instance $Z$. If  $X$ be the exemplar found at training time,
then we displace $Z$ as follows:
\[Z_j =  Z_j + F*(X_j - Z_j)\]
Here, $F$ is the  magnitude controller, discussed above.
The experiments of this work showed that different kinds of data require
different kinds of exemplars.
\bi
\item
For the Seigmund data,
the objective score is the runtime of the test suite of a compiled program.
For this kind of data with numeric objectives,
the exemplar  is the instance in a cluster
with  best objective score.
\item
For the Jureczko data, the objective score
is a boolean that is true if defects have been noted
for that class. 
For this kind of data with discrete objectives,
the data, 
has less differentiation between the objective scores
so we set the exemplar to  
an instance synthesized from the median values of the instances
in that cluster.
\ei

\begin{figure}[!t]
\small 
\[
\begin{array}{r}
\mathrm{project}\\
\mathrm{data}
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; HOW)}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{predictor}(\mathit{test})\\
           \mathrm{\bf if\;}\mathit{before} & <  \mathit{satisfactory}\\
           \mathrm{\bf then}  & \mathrm{{\bf return}\; 0}\\~\\
           \mathrm{\bf else} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} \mathit{after} /  \mathit{before}
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]

@@

  pd & pf & good? & pd & pf & good? & pd & pf\\\cline{1-8}
  50 & 27 &   & 64 & 28 & y & 14 & 1&$\star$\\
  65 & 35 & y & 65 & 25 & y & 0 & -10&$\star$\\
  50 & 30 &   & 62 & 41 &   & 12 & 11\\
  66 & 21 & y & 63 & 18 & y & -3 & -3\\
  45 & 18 &   & 47 & 15 &   & 2 & -3\\
  79 & 61 &   & 77 & 61 &   & -2 & 0\\
  55 & 25 &   & 60 & 27 & y & 5 & 2\\
  51 & 32 &   & 61 & 12 & y & 10 & -20&$\star$\\
 30 & 28 &   & 39 & 33 &   & 9 & 5&$\times$\\
  32 & 6 &   & 30 & 9 &   & -2 & 3&$\times$\\
  41 & 5 &   & 48 & 0 &   & 7 & -5&\times\\
\hline
  \multicolumn{5}{c}{~}    &  & -1 & -4& 25th\\
\multicolumn{6}{r|}{the percentiles of the changes:}    &   5 & 0 & 50th\\
 \multicolumn{5}{c}{~}             & & 10 & 3 & 75th\\
\end{tabular}

@@

\section{When Not to Plan: \\the Problem of Poor Predictors}

@@

For the  Jureczko   data,
the plans generated by HOW advise   how to change static code attributes in order to reduce defects in
Java classes.  Here,
performance can be measured in terms of:
\bi
\item Probability of detection (a.k.a. ``pd'' or recall) which is the percent of faulty classes in
the test data detected
by the {\em predictor}.
\item Probability of false alarm (a.k.a. ``pf'') which is the percent of non-fault
classes that are predicted to be defective.
\ei 
The ``untuned'' columns of \fig{j} shows RandomForest's learned {\em pd,pf}
values generated after learning from the training data, then being applied to the test data.
If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
then only two of our data sets ({\em ivy,ant}) are ``good'' enough.

The ``tuned'' columns of \fig{j} show that we can salvage some of the data sets
using two techniques.
 Pelayo and Dick~\cite{pelayo07} report that defect prediction can be improved via SMOTE~\cite{Chawla2002}; i.e. an over-sampling of minority-class examples.
 Also, Fu et al.~\cite{fu:ase15} report that an optimization method called differential evolution~\cite{storn97}
applied to defect predictors can quickly explore the tuning
options of Random Forest to find better settings for the size of the forest, the termination criteria
for tree generation, etc.
The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these
techniques. For example, in {\em poi}, the recall increased by 10\% while the false alarm rate dropped by 20\%).
However,  as might be expected
the the three bottom rows remain
unsalvageable.

In summary, while we cannot trust predictors from some of our data sets,
we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
(Aside: note that this tuning used optimization and sub-sampling   on   {\em only} the
training instances of \fig{j}
i.e.
no clues from the test set were ever used in this tuning process.)

@@

blah blah data. algorithms. statistics

\section{Releated Work}

Our prior work in generating policies relied upon a Monte Carlo
analysis of an underlying model~\cite{me07f}. This approach had three major problems.
Firstly, enough domain knowledge had to be collected to build that model
and this can take a long time. For example, in our previous work we used three software process models from
the Unviersity of Southern California that were initially
developed and updated for nearly two decades from 1981~\cite{boehm81} to 2000~\cite{boehm00b}.

Secondly, there is the computational cost of running those models.
While

~\cite{}-- which suffered from the problem of (a)~the effort required to collect the domain knowledge to build
problem in certifying the underlying model;
 (b)~the computational problem of running that model enough
times to learn patterns in that data; and (c)~ requrearea used large scale simulations or large

While
our case studies are defect and runtime reduction, HERE is a general
purpose algorithm that could be applied to any table of data with at least
one column whose values can be sorted from ``desired'' to ``undesired''.

{\bf RQ1: Can Instance-based Learning techniques be used to create Planning tools that can better analyse software defect?}

{\bf RQ2: Can lessons learnt from local policy generators be useful for generating solutions that can be used to mitigate defects in software systems?}

{\bf RQ3: Can such instance-based planners be applied to other software engineering paradigms?}
We check if the proposed instance-based planner is applicable to performance prediction.

\section{Motivating Example}
\section{Background}
\section{Instance Based Learning}

Assessing the quality of solutions to a real-world problem can be extremely hard~\cite{menzies2005xomo}. In several software engineering applications, researchers have models that can emulate the problem, for instance there is the COCOMO effort model~\cite[p29-57]{boehm2009software}, the COQUALMO defect model~\cite[p254-268]{boehm2009software}, Madachy’s schedule risk model~\cite[p284-291]{boehm2009software}, to name a few. Using these models it is possible to examine several scenarios in a short period of time, and this can be done in a reproducible manner. However, models aren't always the solution, as we shall see.

There exist several problems where models are hard to obtain, or the input and output are related by complex connections that simply cannot be modeled in a reliable manner, or generation of reliable models take prohibitively long ~\cite{Ludewig2003}. Software defect prediction is an excellent example of such a case. Models that incorporate all the intricate issues that may lead defects in a product is extremely hard to come by. Moreover, it has been shown that models for different regions within the same data can have very different properties \cite{localvsglobal}. This makes it extremely hard for one to design planning systems that are capable of mitigating these defects.

An alternative approach would be to make use of an instance based approach, an example of case based reasoning strategy, in place of the conventional model based approach. Instance based approaches have been used extensively by the effort estimation community, see~\cite{keung2008analogy, 6600685, walkerden1999empirical, shepperd1997estimating, kocaguneli2010use}.

This approach has been proposed as an alternative to closed form mathematical models or other modeling methods such as regression \cite{keung2008analogy}. There are several other reasons for instance based approaches being a useful tool, see~\cite{6600685}. As pointed out by~\cite{walkerden1999empirical} it can be used with partial knowledge of the target project at an early stage which could be a very useful tool in preventing software defects. Instance based approaches are also rather robust in handling cases with sparse samples \cite{1438374}. All these features are desirable and suggest that instance-based approach is a useful adjunct to traditional model based approach.

In this paper the instance based learning principle is used to first cluster the training data and then create contours in the data. These contours can then be used by test cases to infer meaningful information in order to improve it. The entire process is described in greater detail in the following sections. We would like to stress that the purpose of our work is not to promote instance-based planning over model based analytics, rather it is to suggest an alternative when model based approaches cannot be used.

\subsection{Spectral Learning using WHERE}
The algorithm uses WHERE to recursively cluster the input data to identify subsets in the training data that a test instance can learn from. WHERE is a spectral learner which uses the FastMap heuristic to estimate the first principal component. It then recursively partitions the data into two halves along the Median point of the projection on the first principal component, terminating when a half has less than $\sqrt{N}$ items.

\subsection{Planning changes using HERE}
We propose the HERE algorithm as a tool to plan changes in the original test data. Figure~\ref{fig:what} highlights the procedure by which HERE operates. It begins by creating clusters which are generated using the WHERE algorithm discussed above. Following this, a nearest neighbour scheme is applied to identify pairs of nearby clusters. A contour is constructed with these clusters at the vertices to characterize the cluster pairs.

The mutation policy works by projecting the test instance onto the contours and identifying the the contour on which the test instance has the largest scalar projection. The planner then reflects over the vertices of the chosen contour, identifying the \textit{better} vertex among the two. Now, a new instance is generated by \textit{mutating the attributes} of the test instance towards the better vertex and away from the worse vertex. This process is repeated for all the test instances that are considered defective by the defect prediction scheme, discussed below.

\subsection{Feature Weighting} \label{fwt}
The algorithm also allows for a feature weighting and pruning scheme. We have employed a simple (and fast) attribute ranking method where the features are sorted based on their ability to find meaningful splits in the dependent class of the data, see ~\cite{hall03}. Features that can find splits with the \textit{lesser variance} in the dependent variable are \textit{ranked higher} than features that don't.

We reason that the most informative feature must undergo a larger change as compared to other features. Therefore, when feature weighting is employed, the higher ranked features are mutated by a larger extent. Additionally, we have also included information pruning. Here we mutate only the top X\% of the features in each test instance. The remaining features remain unaltered.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{_figs/WHAT-Clusters2.pdf}
\caption{HERE}
\label{fig:whatflow}
\end{figure}

\subsection{Defect Prediction}

To validate the treatments that have been suggested by our planner, we need to have defect predictors that capable of identifying if a certain module may (or may not) have a defect. A recent IEEE TSE paper by Lessmann et al.~\cite{lessmann} compared 21 different learners for software defect prediction: 
\bi
\item
{\em Statistical classifiers:}
Linear    discrimin analysis,
Quadratic discrimin analysis,
Logistic regression,
Naive Bayes,
Bayesian networks,
Least-angle regression,
Relevance vector machine,

\item
{\em Nearest neighbor methods:}
k-nearest neighbor,
K-Star

\item
{\em Neural networks:}
Multi-Layer Perceptron,
Radial bias functions,

\item
{\em Support vector machine-based classifiers:}
Support vector machine,
Lagrangian SVM
Least squares SVM,
Linear programming,
Voted perceptron,

\item
{\em Decision-tree approaches:}
C4.5,
CART,
Alternating DTs
\item
{\em Ensemble methods:}
\textbf{Random Forest},
Logistic Model Tree.
\ei

They concluded that Random Forrest was the best method, CART being the worst.

Random Forest is an ensemble learning scheme that constructs a number of decision trees at the training time, for a test instance it outputs the mode of the classes of individual tree. It's patent from how random forest operates that the prediction will suffer if there is an imbalance in classes during the training. Unfortunately, the data sets explored here do suffer from severe skewness, as highlighted in~\ref{fig:classimb}. A study conducted by Pelayo and Dick~\cite{smote2} inspected this issue. They showed that the SMOTE technique~\cite{smote} can be used to improve recognition of defect-prone modules.

In short, SMOTE works by under-sampling the majority class and oversampling all the minority classes in the training data. We use a similar approach with one minor addition to the original algorithm. In our implementation of SMOTE we have introduced an additional step called \textit{resampling}, wherein we ensure that after we do the over/under sampling, the new training data does not have any of the original rows. The new training data merely resembles the original data. This is done as a precautionary measure, so as not to have HERE and Random Forest train on the same training data. 
%Its worth noting that both HERE and Random Forest require some data to train. In order to prevent the two from using the same data to train, we run SMOTE by  \textit{resampling} the classes. Re-sampling

\begin{figure}[!tb]
  \renewcommand{\baselinestretch}{0.8}\begin{center}
{\scriptsize
\begin{tabular}{l@{~~~}l@{~~~}l@{~~~}l@{~~~}l@{~~~}l@{~~~}l@{~~}}
  \hline
  \rowcolor{lightgray}
  Data & Symbol & Training & Testing & Training Samples& Defective &\% Defective \\\hline

Ant & ant & 1.5, 1.6  &1.7 & 644&124&19.25\\

Camel & cam & 1.2, 1.4 & 1.6 & 1480&361 & 24.39\\

Ivy & ivy & 1.1, 1.4 & 2.0  & 352 & 79 & 22.44\\

Jedit & jed & 4.1, 4.2 & 4.3 & 679 & 127 & 18.70\\

Log4j & log & 1.0, 1.1 & 1.2 & 244 & 71 & 29.09\\

Lucene & luc & 2.0, 2.2 & 2.4 & 442 & 235 & 53.16\\

Poi & poi & 2.0, 2.5 & 3.0 & 699 & 285 & 40.77\\

Synapse & syn & 1.0, 1.1 & 1.2 & 379 & 76 & 20.05\\

Velocity & vel & 1.4, 1.5 & 1.6 & 410& 289 & 70.48\\

Xalan & xal &2.5, 2.6 &2.7 & 1688 & 798 & 47.27\\\hline
\end{tabular}}
\end{center}
\caption{Attributes of the defect datasets}\label{fig:ck}
\end{figure}

@@

\begin{figure}[tbp!]
% \subfloat[\label{oracle}]{\includegraphics[width=\linewidth]{{_figs//Oracle}}\\
\includegraphics[width=\linewidth]{_figs/WHAT}
\caption{Experimental Rig}
\label{fig:rig}

@@

\end{figure}

%involves two key components. Firstly, there is the oracle (Random Forest with SMOTE) which is used to make predictions. Then we have the 
\section{Data sets}

The following section describes the experimental rig and the experiments used to measure the performance of HERE on 10 defect data sets and 6 performance prediction data sets.

\subsection{Defect Data Set}
The data was obtained from the PROMISE repository\footnote{Promise Repository: \url{openscience.us/repo}}. For the defect data we investigated 32 releases from 11 open source Java projects defined by the metrics highlighted in figure ~\ref{fig:ck}: \kw{Apache Ant} (1.5 -- 1.7), \kw{Apache Camel} (1.2 -- 1.6), \kw{Apache Ivy} (1.1 -- 2.0), \kw{JEdit} (4.1 -- 4.3), \kw{Apache Log4j} (1.0 -- 1.2), \kw{Apache Lucene} (2.0 -- 2.2), \kw{PBeans} (1.0 and 2.0), \kw{Apache POI} (2.0 -- 3.0), \kw{Apache Synapse} (1.0 -- 1.2), \kw{Apache Velocity} (1.4 -- 1.6), and \kw{Apache Xalan-Java} (2.5 -- 2.7).

>>>>

 \documentclass[conference]{IEEEtran}
\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 %\usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage[running]{lineno}
\usepackage{program}
\usepackage{cite}
\usepackage{alltt}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\definecolor{lightgray}{gray}{0.975}
\usepackage{fancyvrb}
\usepackage{stfloats}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amsmath}  
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}

\usepackage{color}
\newcommand{\colorrule}[1]{\begingroup\color{#1}\hrule\endgroup}

@@

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=right,
numberstyle=\scriptsize,
stepnumber=1,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\scriptsize\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, Line,bPruning,clister,train, exemplar,nearestLine,dist,displace,geometry,splitAcross2Points,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
firstnumber=last
%
}}{}

@@

Data can be rapidly and recursively divided   as follows.
Find   two   distance instances,  $X,Y$
by picking any instance $W$ at random, then setting $X$ to its most
distant instance, then setting $Y$ to the instance most distant from
$X$~\cite{fastmap}
(this requires only $O(2N)$ comparisons
of $N$ instances).
Next, project each instance $Z$
onto a line that  runs between $X,Y$ using the cosine
rule of \fig{where}b. After that,  split the data at the median $x$ value of all instances and
recurses on each half  (stopping when
one half has less  than $\sqrt{N}$ of the original population). For more details. \fig{howcode}, line~76,

@@

Let   $X,Y$ be two points joined by  a line  length $c$.
A third point, $Z$, has distances  $a=dist(Z,X)$ and
$b=dist(Z,Y)$ to $X,Y$, respectively.
According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
Further, according to Pythagoras' theorem, $Z$ stands at a distance
$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$. For more details, see  \fig{howcode}, line~59.

@@

HERE's feature weighting algorithm
comes from the CART regression tree learner~\cite{Breiman1984}.
It sorts independent variables
 according to how well they can reduce the variance
of a  numeric objective.
If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
then the expected
value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
where $v$ is the standard deviation of the objective.
A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
then split at $\argmin_i w_i$, then recursing into each half.  For more details, see  \fig{howcode}, line~109.

@@

Those same regions can be used for instance-based 
{\em planning}-- but instead of leaving test instances in the regions
found during training, we {\em displace} them towards regions
with better objective scores.
To achieve this, 
during training, HOW clusters data using the procedure of \fig{where}a,
then draws {\tt Line}s between the centroids of  clusters $X,Y$ that are nearest
neighbors.
By convention, we say that the cluster containing $X$ has the  better objective scores. This means we can say that the {\tt Line} is a useful direction for displacing
data; i.e.  away from $Y$ and towards $X$ (for an implementation of such a {\tt Line} class, see \fig{howcode}, line~1).
 During testing,  HOW   finds the  nearest {\tt Line} to every test instance. That  test instance is   then displaced towards the better end
of that line; i.e. towards  $X$ (see \fig{howcode}, line 53).

An important design decision with HOW 
is how to determine the exact {\em direction} and {\em magnitude} of the
displacement.  Following on from the discussion around   ISSUE~\#1,
we note that a design goal of HOW is
{\em minimal displacements} of project data.
To implement minimal displacement:
\bi
\item
HOW sorts  the independent variables
using the procedure of  \fig{where}d. It then only
displaces the    best $B$\%  of that sort, while leaving the rest untouched
(see \fig{howcode}, line 38).
In  the experiments shown below,  we explore   $B \in \{25,50,100\}\%$.
\item
HOW also  limits displacement
to very small regions in the space of the training data (never more that the separation
of the two closest clusters).  HOW's clusterer divides  data into many   small clusters containing
$\sqrt{N}$ of the population (using the {\tt cluster} function, \fig{howcode}, line 76).
HOW   limits displacement to  \mbox{$0 \le F \le c$}, where
$c$ is the distance between the two closest clusters. 
In  the experiments shown below,  we explore
\mbox{$F\in \{0.25, 0.5, 0.75\}$}.
\ei
Turning now to the displacement {\em direction},
our experiments show that it useful to  carefully reflect on exactly where HOW should  push each test instance.
HOW replaces each cluster   with one {\em exemplar} that best
characterizes the data in that cluster. It then displaces the
test data towards that exemplar, as shown in \fig{howcode}, line 19. In summary, 
If  $Z_j$ is an independent variable within the
test instance $Z$ and  $X$ is the exemplar found at training time,
then we displace $Z$ as follows:
\[Z_j =  Z_j + F*(X_j - Z_j)\]
Here, $F$ is the  magnitude controller, discussed above.
The experiments of this work showed that different kinds of data require
different kinds of exemplars.
\bi
\item
For the Seigmund data,
the objective score is the runtime of the test suite of a compiled program.
For this kind of data with numeric objectives,
the exemplar  is the instance in a cluster
with  best objective score.
\item
For the Jureczko data, the objective score
is a boolean that is true if defects have been noted
for that class. 
For this kind of data with discrete objectives,
the data, 
has less differentiation between the objective scores
so we set the exemplar to  
an instance synthesized from the median values of the instances
in that cluster.
\ei

\begin{figure}[!t]
\small 
\[
\begin{array}{r}
\mathrm{project}\\
\mathrm{data}
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; HOW)}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{predictor}(\mathit{test})\\
           \mathrm{\bf if\;}\mathit{before} & <  \mathit{satisfactory}\\
           \mathrm{\bf then}  & \mathrm{{\bf return}\; 0}\\~\\
           \mathrm{\bf else} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} \mathit{before} /  \mathit{after}
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]

@@

  pd & pf & good? & pd & pf & good? & pd & pf\\\cline{1-8}
  50 & 27 &   & 64 & 28 & y & 14 & 1&$\star$\\
  65 & 35 & y & 65 & 25 & y & 0 & -10&$\star$\\
  50 & 30 &   & 62 & 41 &   & 12 & 11\\
  66 & 21 & y & 63 & 18 & y & -3 & -3\\
  45 & 18 &   & 47 & 15 &   & 2 & -3\\
  79 & 61 &   & 77 & 61 &   & -2 & 0\\
  55 & 25 &   & 60 & 27 & y & 5 & 2\\
  51 & 32 &   & 61 & 12 & y & 10 & -20&$\star$\\
 30 & 28 &   & 39 & 33 &   & 9 & 5&$\times$\\
  32 & 6 &   & 30 & 9 &   & -2 & 3&$\times$\\
  41 & 5 &   & 48 & 0 &   & 7 & -5&$\times$\\
\hline
  \multicolumn{5}{c}{~}    &  & -1 & -4& 25th\\
\multicolumn{6}{r|}{the percentiles of the changes:}    &   5 & 0 & 50th\\
 \multicolumn{5}{c}{~}             & & 10 & 3 & 75th\\
\end{tabular}

@@

\section{When Not to Plan: the Problem of Poor Predictors}

@@

For the  Jureczko   data,
the plans generated by HOW advise   how to change static code attributes in order to reduce defects in
Java classes.  Here,
performance can be measured using:
\bi
\item Probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
the test data detected
by the {\em predictor}.
\item Probability of false alarm (a.k.a. ``pf''): the percent of non-fault
classes that are predicted to be defective.
\ei 
The ``untuned'' columns of \fig{j} shows RandomForest's learned {\em pd,pf}
values generated after learning from the training data, then being applied to the test data.
If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
then only two of our data sets ({\em ivy,ant}) are ``good'' enough.

The ``tuned'' columns of \fig{j} show that we can salvage some of the data sets
using two techniques.
 Pelayo and Dick~\cite{pelayo07} report that defect prediction can be improved via SMOTE~\cite{Chawla2002}; i.e. an over-sampling of minority-class examples.
 Also, Fu et al.~\cite{fu:ase15} report that an optimization method called differential evolution~\cite{storn97}
applied to defect predictors can quickly explore the tuning
options of Random Forest to find better settings for the size of the forest, the termination criteria
for tree generation, etc.
The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these
techniques. For example, in {\em poi}, the recall increased by 10\% while the false alarm rate dropped by 20\%.
However,  as might be expected
the data sets in the  three bottom rows remain
unsalvageable.

In summary, while we cannot trust predictors from some of our defect data sets,
we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
(Aside: note that this tuning used optimization and sub-sampling   on   {\em only} the
training instances of \fig{j}
i.e.
no clues from the test set were ever used in this tuning process.)

@@

To assess HOW, we applied its plans to all the Seigmund data, and the five ``good''
  Jureczko   data sets.
  The training and tests used for Jurecko were display in \fig{j}. For Siegmind,
  25 times, 
  we randomized the order of the data, trained on the first half, then tested on the second half.

  \fig{ck} shows the attributes in the Jureczko data.
  The Seigmund data sets are far more varied (every configuration option in the Makefiles
  for BerkeleyDB-- the Java  and  Windows  versions,  Apache,  SQLite,  LLVM,and x264).
  \fig{cpm} lists the size of Seigmund attribute sets.

@@

%\begin{figure}[tbp!]
% \subfloat[\label{oracle}]{\includegraphics[width=\linewidth]{{_figs//Oracle}}\\
%\includegraphics[width=\linewidth]{_figs/WHAT}
%\caption{Experimental Rig}
%\label{fig:rig}

@@

%\end{figure}

%involves two key components. Firstly, there is the oracle (Random Forest with SMOTE) which is used to make predictions. Then we have the

We  assess the performance of HOW  by measuring the gain, given by $\frac{Before}{After}$. Since
our goal is to reduce all our objective scores (defects and runtimes of the compiled systems).
the gain takes a value larger than 1 if the run times have been reduced, a value of 1 if there is no change, and a value less than 1 if run times have increased after applying the recommended changes.

Second, we summarizes all our runs using some visualizations and some statistical tests.
In those statistical tests, in  order to avoid any parametric bias,
we use the Cliffs Delta score to measure the probability that the number of bugs in test data before applying the planner is larger than after doing so. In other words, we use the delta score to measure the ability of the planner to effective reduce the number of defects. Given the untreated test instance labeled \textit{Before} of length \textit{M} and the treated instances labeled \textit{After}, the delta score is obtained as follows:
\begin{equation}
\delta = \frac{\#(Before<After) - \#(Before>After)}{M^2}
\label{eq:cliffs}
\end{equation}

A $\delta$ of  0 implies no change at all.

For the performance prediction data set on the other hand, Cliff's delta measure is not directly applicable.

In addition to the above, we rank the different various of the planning scheme to identify the best approach. We make use of the Scott-Knott procedure, recommended by Mittas \& Angelis in their 2013 IEEE TSE paper~\cite{sk}, to compute the ranks. It works as follows. A list of treatments \textit{l} is sorted by their median score. The list \textit{l} is then split into sub-lists $m, n$ in order to maximize the expected value of the differences in the observed performance before and after division. A statistical hypothesis test \textit{H} is applied on the splits \textit{m, n} to check if they are statistically different. If so, Skott-Knott then recurses on each division.
